Capstone Movielens Report

### 2025-11-11 reports/capstone-movielens.site/02-0.methods.Rmd
Course textbook chapter number changed: 25 -> 24.

We will split the `edx` dataset into a training set, which we will use to build and train our models, and a test set in which we will compute the accuracy of our predictions, the way described in [Section 25.1 *Case study: recommendation systems*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-recommendation-systems) of the _Course Textbook (New Edition)_ mentioned above[@IDS2_25-1].We will also use the _5-Fold Cross Validation_ method as described in [Section 31.6 *Cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#cross-validation) of the _Course Textbook (New Edition)_[@IDS2_31-6]. 
To prepare datasets for processing, we will use the following functions (described in detail in the [Initializing Input Datasets] section of the [Appendix A: Support Functions] of *this Report*), specifically designed for these operations:

>>>>

We will split the `edx` dataset into a training set, which we will use to build and train our models, and a test set in which we will compute the accuracy of our predictions, the way described in [Section 24.1 *Case study: recommendation systems*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-recommendation-systems) of the _Course Textbook (New Edition)_ mentioned above[@IDS2_25-1].We will also use the _5-Fold Cross Validation_ method as described in [Section 31.6 *Cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#cross-validation) of the _Course Textbook (New Edition)_[@IDS2_31-6]. 
To prepare datasets for processing, we will use the following functions (described in detail in the [Initializing Input Datasets] section of the [Appendix A: Support Functions] of *this Report*), specifically designed for these operations:

### 2025-11-11 reports/capstone-movielens.site/02.4.UE-model.Rmd
Course textbook chapter number changed: 25 -> 24.

To improve our model let's now take into consideration user effects as explained in [Section 25.1 *Case study: recommendation systems / User effects*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#user-effects) of the *Course Textbook (New Edition)*.

>>>>

To improve our model let's now take into consideration user effects as explained in [Section 24.1 *Case study: recommendation systems / User effects*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#user-effects) of the *Course Textbook (New Edition)*.

### 2025-11-11 reports/capstone-movielens.site/02.5.UME-model.Rmd
Course textbook chapter number changed: 25 -> 24.

::: {#textbook.25.1.movie_effects .sidebar }
In [Section 25.1 *Case study: recommendation systems / Movie effects*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#movie-effects) of the *Course Textbook (New Edition)*, the author draws our attention to the fact that some movies are generally rated higher than others[@IDS2_25-1]. 
:::

@@

 [Section 25.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_ explains why and how we should use _Penalized least squares_ to improve our predictions. The author also explains that the general idea of penalized regression is to control the total variability of the movie effects[@IDS2_25-1]: 
$$
\sum_{j=1}^{n} \beta_j^2
$$ 
Specifically, instead of minimizing the least squares equation, we minimize an equation that adds a penalty:

>>>>

::: {#textbook.25.1.movie_effects .sidebar }
In [Section 24.1 *Case study: recommendation systems / Movie effects*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#movie-effects) of the *Course Textbook (New Edition)*, the author draws our attention to the fact that some movies are generally rated higher than others[@IDS2_25-1]. 
:::

@@

 [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_ explains why and how we should use _Penalized least squares_ to improve our predictions. The author also explains that the general idea of penalized regression is to control the total variability of the movie effects[@IDS2_25-1]: 
$$
\sum_{j=1}^{n} \beta_j^2
$$ 
Specifically, instead of minimizing the least squares equation, we minimize an equation that adds a penalty:

### 2025-11-11 reports/capstone-movielens.site/05.appendix.Rmd
Course textbook chapter number changed: 25 -> 24.

::: { #func.clamp.details .sidebar }
##### Details
\
As explained in [Section 25.1 *Case study: recommendation systems / User effects*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#user-effects) of the _Course Textbook_ we know ratings cannot be below 0.5 or above 5. For this reason, we need a function described in that section[@IDS2_25-1]:

@@

Calculates *Mean Squared Error* as described in [Section 25.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_25-1].

@@

::: { #func.mse.args.r .sidebar }
  - **r:** Numeric vector representing a set of residuals as described in [Section 25.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_25-1].

@@

::: { #func.mse_cv.args.r_list .sidebar }
  - **r_list:** List of numeric vectors, each representing a set of residuals as described in [Section 25.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_25-1], calculated in `K`th iteration using the _K-Fold Cross Validation_ method.

@@

Calculates *Root Mean Squared Error* as described in [Section 25.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_25-1].

@@

::: { #func.rmse.args.r .sidebar }
  - **r:** Numeric vector representing a set of residuals as described in [Section 25.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_25-1].

@@

We call this function internally from the functions that compute the _penalized estimates_  as described in section [Section 25.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_25-1].

@@

The *Regularization technique* (described in section [Section 25.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_25-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the section [Regularization Functions].
:::

@@

Note that titles cannot be considered unique, so we cannot use them as IDs[@IDS2_25-1].

>>>>

::: { #func.clamp.details .sidebar }
##### Details
\
As explained in [Section 24.1 *Case study: recommendation systems / User effects*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#user-effects) of the _Course Textbook_ we know ratings cannot be below 0.5 or above 5. For this reason, we need a function described in that section[@IDS2_24-1]:

@@

Calculates *Mean Squared Error* as described in [Section 24.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_24-1].

@@

::: { #func.mse.args.r .sidebar }
  - **r:** Numeric vector representing a set of residuals as described in [Section 24.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_24-1].

@@

::: { #func.mse_cv.args.r_list .sidebar }
  - **r_list:** List of numeric vectors, each representing a set of residuals as described in [Section 24.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_24-1], calculated in `K`th iteration using the _K-Fold Cross Validation_ method.

@@

Calculates *Root Mean Squared Error* as described in [Section 24.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_24-1].

@@

::: { #func.rmse.args.r .sidebar }
  - **r:** Numeric vector representing a set of residuals as described in [Section 24.1 *Case study: recommendation systems / Loss function*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook (New Edition)_[@IDS2_24-1].

@@

We call this function internally from the functions that compute the _penalized estimates_  as described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1].

@@

The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the section [Regularization Functions].
:::

@@

Note that titles cannot be considered unique, so we cannot use them as IDs[@IDS2_24-1].
